### 2017-05-11 acknowledgments.tex

\section*{Acknowledgments}
\label{sec:acknowledgments}

% Blind review. 
% Procurar no texto por Institution 2 = University of São Paulo.
% Retornar o texto abaixo com o agradecimento para as fundações.

The authors would like to thank Students who participated in this exercise, CNPq, CAPES (BEX 3401/15-4, CAPES-COFECUB Project No. 828/15), São Paulo Research Foundation (FAPESP) (procs. \#2012/23300-7) for the fellowships and the Araucária Foundation, Department of Science, Technology and High Education of State Government of Paraná (SETI-PR) and State Government of Paraná for the financial support.

% The authors would like to thank Students who participated in this exercise and financial support agencies.