### 2017-07-13 sections/02-background.tex
Re-organizing and fixing the used template

\section{Background}
\label{sec:background}

TO-DO ...