### 2017-07-13 sections/04-researchdesign.tex

\section{Research Design}
\label{sec:researchdesing}

The motivation to write this paper comes from experiences on teaching. During the classes of Distributed Systems and Parallel Programming, a graduate course at University of São Paulo...