### 2017-07-13 seps-omp-paper.tex

\include{configurations}

\begin{document}
%---
\title{OpenMP is Not as Easy as It Appears, \\
      but It is the Easiest One}
%---
\author{
    \IEEEauthorblockN{Pedro Bruel\IEEEauthorrefmark{1},
                      Paulo Meirelles\IEEEauthorrefmark{1} and
                      Alfredo Goldman\IEEEauthorrefmark{1}}
    \IEEEauthorblockA{\IEEEauthorrefmark{1}Software Systems Laboratory -- LSS\\
    Institute of Mathematics and Statistics -- IME\\
    University of São Paulo -- USP\\
    São Paulo/SP, Brazil\\
    \{phrb,paulormm,gold\}@ime.usp.br}
}
%---
\maketitle
%---
\input{sections/00-abstract}
%---
\IEEEpeerreviewmaketitle
%---
\input{sections/01-introduction}
\input{sections/02-background}
\input{sections/03-relatedwork}
\input{sections/04-researchdesign}
\input{sections/05-dataanalysis}
\input{sections/06-lessons}
\input{sections/07-conclusions}
%---
\input{sections/acknowledgments}
%---
\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,seps-omp-paper}
\end{document}