### 2017-07-21 sections/04-researchdesign.tex
Re-defining initial research questions

\section{Research Design}
\label{sec:researchdesing}

The motivation to write this paper comes from experiences on teaching. During the classes of Distributed Systems and Parallel Programming, a graduate course at University of São Paulo...

\begin{description}
\item[RQ1:] \textit{Which API is easier to learn and use according to the student perceptions?}
%
...
\item[RQ2:] \textit{Which API is better to improve the performance according to the student perceptions?}
%
...

\end{description}