### 2017-07-29 sections/05-results.tex

\section{Results}
\label{sec:results}

\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.85\columnwidth]{likert_questions}
    \caption{Student responses to Likert-scale questions}
    \label{fig:}
\end{figure}

\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.85\columnwidth]{learning_mean_difficulty}
    \caption{Student mean difficulty to learn the APIs}
    \label{fig:}
\end{figure}

\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.85\columnwidth]{comparisons}
    \caption{Student responses on comparisons of APIs, and the performance of their usage of the APIs}
    \label{fig:}
\end{figure}