### 2017-08-01 sections/02-background.tex
Add text to responses section, improve figures

\section{Related Work}
\label{sec:relatedwork}

\todo[inline,color=cyan,author=Pedro]{Cite our previous paper, making it clear that we are the authors}
\todo[inline,color=cyan,author=Pedro]{Describe related work}
\todo[inline,color=cyan,author=Pedro]{Describe related work since last paper}