### 2017-08-01 sections/02-background.tex
Change related work section name.

\section{Background}
\label{sec:background}

\todo[inline,color=cyan,author=Pedro]{Cite our previous paper, making it clear that we are the authors}
\todo[inline,color=cyan,author=Pedro]{Describe related work}
\todo[inline,color=cyan,author=Pedro]{Describe related work since last paper}