### 2017-08-01 sections/03-apis.tex
Improve responses section.

\section{OpenMP \& Pthreads}
\label{sec:apis}

\todo[inline,color=cyan,author=Pedro]{Provide brief descriptions of the APIs}

\subsection{Pthreads}

\todo[inline,color=cyan,author=Pedro]{Describe Pthreads}

\subsection{OpenMP}

\todo[inline,color=cyan,author=Pedro]{Describe OpenMP}