### 2017-08-01 sections/04-research-design.tex

\section{Research Design}
\label{sec:researchdesign}

\todo[inline,color=cyan,author=Pedro]{Describe research design}

\subsection{The Concurrent, Parallel and Distributed Computing Course}

\todo[inline,color=cyan,author=Pedro]{Describe the course}

\subsection{Questionnaire}

\todo[inline,color=cyan,author=Pedro]{Describe the questionnaire}
\todo[inline,color=cyan,author=Pedro]{Add reference for questionnaires from Herez}

\subsection{Research Questions}

\todo[inline,color=cyan,author=Pedro]{Improve RQ discussion and descriptions}

According to the students' perception, which API:

\begin{description}
    \item[RQ1:] \textit{Was easier to learn and use?}
    \item[RQ2:] \textit{Improved performance the most?}
\end{description}

After evaluation of the students' performance improvements
in the assignment:

\begin{description}
    \item[RQ3:] \textit{Did the students' performance improvements match their
        perception?}
\end{description}