### 2017-08-07 sections/01-introduction.tex

\section{Introduction}
\label{sec:introduction}

\todo[inline,color=cyan,author=Pedro]{Write new introduction}

...

In this work we have gathered data from assignments made by graduate students
for the \emph{Introduction to Parallel and Distributed Computing} course. Among
other exercises, the students were asked to ...

\todo[inline,color=cyan,author=Pedro]{Describe paper structure}