### 2017-08-08 sections/00-abstract.tex

\begin{abstract}
\todo[inline,color=cyan,author=Pedro]{Summarize results}
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%

\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010147.10010169</concept_id>
  <concept_desc>Computing methodologies~Parallel computing methodologies</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010147.10010919</concept_id>
  <concept_desc>Computing methodologies~Distributed computing methodologies</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computing methodologies~Parallel computing methodologies}
\ccsdesc[500]{Computing methodologies~Distributed computing methodologies}

\keywords{Parallel and Distributed Computing, OpenMP, Pthreads}