### 2017-08-08 sections/00-abstract.tex
Add text for the abstract, introduction and conclusion.

\begin{abstract}

    Selecting the most effective tools and languages for teaching parallel and
    distributed computing presents a challenge for teachers and students. Often
    a high-level solution seems easier, but it comes with the cost of hiding
    underlying concepts and problems. We present a comparison between the
    \textit{Pthreads} and \textit{OpenMP} APIs considering the experiences and
    performance results of undergraduate and graduate Computer Science students
    from the University of São Paulo.
    We show that students were able to achieve better performance results using
    \textit{Pthreads} than \textit{OpenMP}, despite their perception that
    \textit{Pthreads} is harder to learn and use.

\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%

\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010147.10010169</concept_id>
  <concept_desc>Computing methodologies~Parallel computing methodologies</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010147.10010919</concept_id>
  <concept_desc>Computing methodologies~Distributed computing methodologies</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computing methodologies~Parallel computing methodologies}
\ccsdesc[500]{Computing methodologies~Distributed computing methodologies}

\keywords{Parallel and Distributed Computing, OpenMP, Pthreads}