### 2017-10-16 sections/05-responses.tex
Change colors to grayscale.

\section{Students' Background \& Responses}
\label{sec:responses}

In this section we present the background of our students in the technologies
relevant to the course and to the assignment. Then we present the responses
given by the students to our questionnaire. The questions are listed in Tables
\ref{tab:likert} and \ref{tab:comparisons}.

%\todo[inline,color=cyan,author=Pedro]{Improve figure colorscheme}

\subsection{Background}

To gauge the proficiency and knowledge of our students we elaborated a
five-point scale from \textit{Do not Know} $(1)$ to \textit{Expert or
Proficient} $(5)$.  We asked the students to express their proficiency level on
this scale, and their responses are shown in Figure \ref{fig:background}.

\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.85\columnwidth]{background_questions}
    \caption{Student knowledge \textit{before} the assignment}
    \label{fig:background}
\end{figure}

It is common for Computer Science introductory and advanced courses in the
graduate and undergraduate programs of the University of São Paulo to be given
in programming languages such as C, Java and Python. Figure
\ref{fig:background} shows that students reported higher proficiency in those
languages. The C language was the one with which students reported most
proficiency.

\begin{table}[htpb]
    \centering
    \begin{tabular}{@{}p{0.9\columnwidth}p{0.08\columnwidth}@{}}
        \toprule
        \multicolumn{1}{c}{\scriptsize{Have you$\dots$}} & \textnumero \\ \midrule
        \scriptsize{Had contact with parallel and distributed concepts before?} & $(1)$ \\
        \scriptsize{Had contact with the APIs before?} & $(2)$ \\
        \scriptsize{Learned from the OpenMP class?} & $(3)$ \\
        \scriptsize{Learned from the Pthreads class?} & $(4)$ \\ \bottomrule
    \end{tabular}
    \caption{Questions for Figure \ref{fig:classes}}
    \label{tab:classes}
    \vspace{-0.6cm}
\end{table}

\begin{figure}[htpb]
    \vspace{-0.7em}
    \centering
    \includegraphics[width=0.85\columnwidth]{classes_questions}
    \caption{Students' responses to questions in Table \ref{tab:classes}}
    \label{fig:classes}
    \vspace{-1em}
\end{figure}

We asked the students to evaluate their proficiency in \textit{OpenMP} and
\textit{Pthreads}, \textit{before} completing the assignment, with two sets of
questions. The first set used the same five-point scale as the other background
questions and is also shown in Figure \ref{fig:background}. The students
reported larger proficiency in \textit{Pthreads} than in
\textit{OpenMP}, and the majority of the students reported they did not know
\textit{OpenMP} before the course.

The second set of questions regarding student proficiency with the APIs
comprised the 4 \textit{Yes or No} questions listed in Table \ref{tab:classes}.
This set of questions also asked the students to evaluate whether the classes
on each subject helped their understanding and learning of the technologies.
The student responses for this set are shown in Figure \ref{fig:classes}.  The
majority of the students have had previous contact with the concepts related to
concurrent, parallel and distributed programming, but the majority also have
not had contact or experience with \textit{OpenMP} or \textit{Pthreads}.  Most
students reported that the classes on each API helped their learning.

\subsection{Responses}

We asked the students to answer 3 sets of questions regarding their
experience with \textit{OpenMP} and \textit{Pthreads}, \textit{after}
completing the assignment.
The first set of questions used a scale defined by us with 5 levels to measure the relationship
of the students with the two APIs in the context of the assignment. This kind of scale enables respondents to express their agreement
or disagreement with specific statements~\cite{likert1932technique}. Respondents choose between five
options: \textit{Strongly Agree} (SA), \textit{Agree} (A), \textit{Neutral}
(N), \textit{Disagree} (D) and \textit{Strongly Disagree}.

The questions of the first set are listed in Table \ref{tab:likert}. The
students expressed their agreement or disagreement with statements related to
the difficulty of parallelizing the first loop with independent iterations,
parallelizing the two nested loops with independent iterations and improving
the performance of the sequential code. Each question was repeated for
\textit{OpenMP} and \textit{Pthreads}, in a total of $6$ questions.

Figure \ref{fig:likert} shows a visualization of the students' responses.  The
height of each stacked sub-column represents the percentage of respondents that
selected each of the correspondent scale agreement levels, which are
represented by different colors.

The percentages of \textit{Strongly Agree} (SA) and \textit{Agree} (A)
responses are larger for questions $(1)$, $(3)$ and $(5)$, which
are related to \textit{OpenMP}.  Likewise, the percentages of \textit{Disagree}
(D) responses are larger for questions $(2)$, $(4)$ and $(6)$,
which are related to \textit{Pthreads}. These responses lead to the conclusion
that \textbf{the perception of the students was that it was easier to use
\textit{OpenMP} than \textit{Pthreads}} to parallelize and improve the
performance of the sequential code provided in the assignment.

\begin{table}[htpb]
    \centering
    \begin{tabular}{@{}p{0.7\columnwidth}p{0.1\columnwidth}p{0.05\columnwidth}@{}}
        \toprule
        \multicolumn{2}{c}{\scriptsize{It is easy to$\dots$}} & \textnumero \\ \midrule
        \multirow{2}{*}{\parbox{0.7\columnwidth}{\scriptsize{Parallelize loops with independent iterations using:}}} & \scriptsize{OpenMP} & $(1)$ \\
        & \scriptsize{Pthreads} & $(2)$ \\
        \addlinespace{}
        \multirow{2}{*}{\parbox{0.7\columnwidth}{\scriptsize{Parallelize nested loops with independent iterations using:}}} & \scriptsize{OpenMP} & $(3)$ \\
        &  \scriptsize{Pthreads} & $(4)$ \\
        \addlinespace{}
        \multirow{2}{*}{\parbox{0.7\columnwidth}{\scriptsize{Improve the performance of sequential code using:}}} & \scriptsize{OpenMP} & $(5)$  \\
        &  \scriptsize{Pthreads} & $(6)$ \\ \bottomrule
    \end{tabular}
    \caption{Scale questions for Figure \ref{fig:likert}}
    \label{tab:likert}
    \vspace{-0.6em}
\end{table}

\begin{figure}[htpb]
    \vspace{-2em}
    \centering
    \includegraphics[width=0.85\columnwidth]{likert_questions}
    \caption{Students' responses to the scale questions in Table \ref{tab:likert}}
    \label{fig:likert}
    \vspace{-1em}
\end{figure}

The second set of questions used a five-point scale to measure the difficulty
the students had with the learning process of the two APIs for the assignment.
Respondents expressed their difficulty to learn \textit{OpenMP} and
\textit{Pthreads} in a scale of $1$ to $5$, choosing a point between \textit{No
Difficulty} $(1)$ and \textit{Much Difficulty} $(5)$.

Figure \ref{fig:learning} shows a visualization of the students' responses.
The height of each stacked sub-column represents the percentage of respondents
that selected each of the correspondent difficulty levels, which are
represented by different colors.

\begin{figure}[htpb]
    \centering
    \includegraphics[width=0.85\columnwidth]{learning_difficulty_questions}
    \caption{Student difficulty to learn the APIs in a five-point scale}
    \label{fig:learning}
\end{figure}

The percentages of \textit{No Difficulty} $(1)$ and difficulty level $(2)$
responses are larger for learning \textit{OpenMP}. Likewise, the
percentages of difficulty level $(4)$ and \textit{Much Difficulty} $(5)$
responses are larger for learning \textit{Pthreads}.  We derive
from these responses the conclusion that \textbf{the students perceived less
difficulty to learn \textit{OpenMP} than \textit{Pthreads}} for the purposes of
the assignment.

The final set of questions regarding the usage and learning of \textit{OpenMP}
and \textit{Pthreads} in the context of the first assignment presented
comparison questions between the two APIs. Students chose which API they
perceived was the best answer to the questions in Table \ref{tab:comparisons}.

Responses to question $(13)$ in Table \ref{tab:comparisons} were not obtained
directly from the students. We analysed each assignment submission and
evaluated the performance of the each solution across the specified
experimental settings.

Figure \ref{fig:comparisons} shows a visualization of the students' responses.
The height of each stacked sub-column represents the percentage of respondents
that selected each API, which are represented by different colors.

The percentage of \textit{OpenMP} selections is larger for all
questions, except for questions $(7)$ and $(13)$. The responses for questions
$(7)$ -- $(12)$ strengthen the previous conclusion that the perception of the
students was that it was easier to use \textit{OpenMP} than \textit{Pthreads}
in the context of the assignment. The performance results that we obtained from
analysing the students' submissions challenge the perception of the students.
The responses to question $(13)$, extracted by our analysis, show that
\textbf{the students achieved better performance when using \textit{Pthreads}
than \textit{OpenMP}}.

\begin{table}[htpb]
    \centering
    \begin{tabular}{@{}p{0.9\columnwidth}p{0.08\columnwidth}@{}}
        \toprule
        \multicolumn{1}{c}{\scriptsize{Which technology$\dots$}} & \textnumero \\ \midrule
        \scriptsize{Was more difficult to learn?} & $(7)$ \\
        \scriptsize{Was simpler to use?} & $(8)$ \\
        \scriptsize{Was easier to use?} & $(9)$ \\
        \scriptsize{Was easier to use for parallelizing loops with independent iterations?} & $(10)$ \\
        \scriptsize{Was easier to use for parallelizing nested loops with independent iterations?} & $(11)$  \\
        \scriptsize{Was easier to use for improving the performance of a sequential program?} & $(12)$  \\
        \scriptsize{Had the best performance on your assignment?} & $(13)$ \\ \bottomrule
    \end{tabular}
    \caption{Questions for Figure \ref{fig:comparisons}}
    \label{tab:comparisons}
    \vspace{-0.6cm}
\end{table}

\begin{figure}[htpb]
    \vspace{-.7em}
    \centering
    \includegraphics[width=0.85\columnwidth]{comparisons}
    \caption{Students' responses to the questions in Table \ref{tab:comparisons}}
    \label{fig:comparisons}
    \vspace{-1em}
\end{figure}